Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-04-05 paper.tex
updated text

{\it IC -- dust and starlight}\\ Galaxy clusters are typically
characterized by the hot gas in the ICM and the collection of
gravitationally bound galaxies. The emission in the IR and UV
wavelengths emerge from dust and starlight in both the galaxies and
the ICM (e.g. \cite{2006ApJ...648L..29P} and
\cite{2009MNRAS.399.1694G}). Much of the radiation leaks out from the
galaxy into the ICM, which give rise to similar spectral distributions
for the galaxies and the ICM in these wavelengths. Hence, we use the
accurately measured light from galactic dust and starlight for the ISM
to model the ICM. We characterize the spectra through a fit to the
galactic spectra presented in \cite{2006ApJ...648L..29P}. The spatial
distribution of the SD light in clusters is extracted from a stacked
emission analysis of Sloan Digital Sky Survey (SDSS) data at the
redshift $\sim 0.25$ \cite{2005MNRAS.358..949Z}. Finally, we fix the
normalization for the light from stars and dust in an galaxy cluster
individually using using the IR to X-ray luminosity scaling relation
found in \cite{2008A&A...490..547G}. The X-ray luminosities for the
clusters used in this work are listed in the extended HIghest X-ray
FLUx Galaxy Cluster Sample (HIFLUGCS) \cite{2002ApJ...567..716R} from
the Rosat all-sky survey. Note that we have added the Virgo cluster to
the sample that we nevertheless refer to as extended HIFLUGCS
catalogue in the following.

{\it Spatial distribution -- stars and dust}\\ Our goal in this
section is to derive a simple spatial model for the distribution of SD
in galaxy clusters. For this reason we use stacked cluster
observations by Sloan in the $r$-band that trace the stars. We assume
that the dust trace the stars in the clusters, which should be
accurate to the order of magnitude estimate that we are interested in
here. The stacked surface brightness profiles in
\cite{2005MNRAS.358..949Z} are measured at redshift $z \sim 0.25$ with
an average mass of the clusters of $4.0\times10^{13}\,\msun$. The
profiles are build up out of three components; diffuse intracluster
light (ICL), galaxies, and the brightest cluster galaxy (BCG) in the
center of clusters. However, the total volume overlap of the ISM to
the cluster is very small compared to the ICL- and BCG overlap with
the cluster, thus the relative contribution of the galaxies to the IC
emission is suppressed. To remove this bias we cut out the galaxies
that have been smoothed over the cluster in the stacked analysis. In
Fig.~\ref{fig:SD_spatial} we show the SDSS stacked brightness profiles
as well as the fitted profiles\footnote{The measured brightness is
  converted into units of $\rmn{erg}\,\rmn{s}^{-1}$ using
  \cite{2010...book} $S(\rmn{mag}\,''^{-2}) =
  \mathcal{M}_\odot+21.572-2.5\rmn{log}_{10}S(L_\odot\,\rmn{pc}^{-2})$,
  where the sun's absolute magnitude in the $r$-band is given by
  $\mathcal{M}_\odot=27.1$ \cite{2000asqu.book..339L} and the
  luminosity of the sun by $L_\odot=3.85\,10^{33}
  \rmn{erg}\,\rmn{s}^{-1}$.}. Our benchmark spatial model are show
with the solid black line, and include the contribution from the ICL
and BCG.

@@

In Fig.~\ref{fig:SD_spectra} we show the spectral distribution of the
light from SD together with the CMB black-body distribution in a
galaxy cluster, both at the radius $r=0.071\rvir$ where the smooth
energy density of the SD light (see Fig.~\ref{fig:SD_Edens}) equals
the energy density of the CMB. We have renormalized both the spectral
data and the fitted spectra using the luminosity from stars and dust
individually (see below for further details about the
renormalization).
\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/fit.porter.v2.eps}
\caption{\it Spectral dependence of radiation fields in a cluster of
  galaxies. The black dotted line in the left peak show the spectrum
  of CMB photons using a black body with a temperature of
  $2.73$~K. The crosses in the middle and right peaks represent the
  measured spectra from stars and dust (SD), respectively, and are
  derived in \cite{2006ApJ...648L..29P} for a galaxy. We normalize the
  individual SD spectrum separately using the observed luminosity from
  SD in clusters. The SD luminosity is related to the cluster mass
  through Eq, where we use have used the mass
  $\mvir=4.0\times10^{13}\msun$ in this figure. Finally we renormalize
  the SD spectra to the radius $r=0.071\rvir$, where the smooth energy
  density of the SD light (see Fig.~\ref{fig:SD_Edens}) equals the
  energy density of the CMB. The red dashed lines show the fitted SD
  spectral model.}
 \label{fig:SD_spectra}
\end{figure}
We choose to fit the spectrum of the dust that peaks at about
$10^{-2}\,\ev$ with a double power-law, and the slightly wider
spectral distribution of the stars that peaks at about $1\,\ev$ with a
tipple power-law. We find the best fit spectra for the SD in a
cluster through:
\begin{eqnarray}
  \eph^2\frac{\dd^2 N_\ph}{\dd \eph \dd V}
  &\equiv& u_\sd(\eph) =  \sum_i N_i(\lx) u_i^\gal(\eph)\,,\nonumber \\ 
\rmn{where}\quad i&=&\{\rmn{stars,dust}\} \qquad \rmn{and}\\ 
% &=&  N_\dust(\lx) u_\dust^\gal(\eph) + N_\stars(\lx) u_\stars^\gal(\eph)  \nonumber \\ 
%  \\
  u_\stars^\gal(\eph) &=& \frac{23\,\rmn{eV}}{\rmn{cm}^3} 
  \left(\frac{1.23\,\rmn{eV}}{\eph}\right)^{1.9} \nonumber \\
  &\times&\left[1+\left(\frac{2.04\,\rmn{eV}}{\eph}\right)^{20}\right]
  ^{-\frac{1.9}{20}}\nonumber \\
  &\times& \left[1+\left(\frac{0.78\,\rmn{eV}}{\eph}\right)^{20}\right]^{-\frac{2.6}{20}}\,, \\
  u_\dust^\gal(\eph) &=& 
  \frac{40\,\rmn{eV}}{\rmn{cm}^3} 
  \left(\frac{0.0144\,\rmn{eV}}{\eph}\right)^{4.9}\nonumber \\
  &\times& \left[1+\left(\frac{0.0144\,\rmn{eV}}{\eph}\right)^{1.9}\right]^{-4.9}\,.
\end{eqnarray}

{\it Normalization}\\ HARD TO READ FIRST PART The spatial and spectral
distribution functions of the light from SD in a cluster are both
normalized with a constant determined from the luminosity in each
respective wavelength. The specific energy density is given by
\begin{eqnarray} 
u_\sd(\eph, r) &=& \frac{j(r)}{j(0)}\,u_\sd(\eph) \nonumber\\
&=& \frac{j(r)}{j(0)}\,\sum_i N_i(\lx) u_i^\gal(\eph)\,,
\end{eqnarray}
where we fix the unitless normalization $N_i(\lx)$ using the total
energy of the light in SD ($E_{i,\vir}$) within $\rvir$:
\begin{eqnarray} 
  E_{i,\vir} &=& L_i \frac{\rvir}{c} \nonumber \\
  &=&N_i(\lx)\int_{\rvir} \int_i \,\frac{j(r)}{j(0)} 
  \frac{u_i^\gal(\eph)}{\eph}\,\dd V\dd \eph\,,\nonumber \\
 \rmn{for}\quad i&=&\{\rmn{stars,dust}\}\,.
\label{eq:E_SD}
\end{eqnarray}
Here we approximate the total energy of the photons within $\rvir$
with the luminosity $L_i$ \cite{2008A&A...490..547G} times the typical
timescale it takes for a photon to propagate through the cluster (we
assume that the cluster is optically thin). Once we derived the SD
luminosities for a well representative cluster, we use a IR to X-ray
scaling relation to estimate the luminosity from an arbitrary cluster:
\begin{equation}
L_\ir=1.05\times10^{45}\,\rmn{erg\,s}^{-1}\,
\left(\frac{\lx}{10^{44}\rmn{erg\,s}^{-1}}\right)^{0.53}\,,
\label{eq:SD-X}
\end{equation}
where $\lx$ is the luminosity of the X-ray emitting gas within
$\rvir$.

@@

{\it Energy density}\\ The spatial distribution of the SD energy
density is an important quantity since it impacts the steady state
electron and positron spectra through cooling. A higher density of SD
in a cluster suppresses the spectra of the upscattering particles. With
the formalism presented above, we can now derive the energy density
from starlight and dust in a galaxy cluster through
\begin{eqnarray}
\label{eq:U_SD}
u_\sd(r) &=& \int \dd \eph \frac{\dd^2 N_\ph}{\dd \eph \dd V}\,\eph
=\int \dd \eph \frac{u_\sd(\eph, r)}{\eph}
\nonumber \\
&=&  \frac{j(r)}{j(0)} \int \dd \eph \sum_i 
N_i(\lx) u_i^\gal(\eph)\,. \nonumber \\
\end{eqnarray}
In Fig.~\ref{fig:SD_Edens} we compare the energy densities from
different radiation fields in a galaxy cluster with the mass
$\mvir=4.0\times10^{13}\msun$. We use two different profiles for the
SD energy density, where the total profile includes the contribution
from the ICL, the BCG and all the galaxies, while the galaxies are
excluded in the smooth profile. We find that even for a cluster with a
relative small mass the energy density of the SD components dominates
over the CMB and the magnetic fields (with a central B field of 3~$\mu
G$) within about 10\% of $\rvir$. Outside this radius the CMB is
dominating the energy density of the cluster, hence the magnetic
fields are always subdominant in the cluster and can safely be removed
from cluster IC calculations for a reasonable central density and
spectral index. Although, in this work we keep the contribution from
B-fields to the total energy density for consistency.

@@

\section{Spectral gamma-ray distribution}
\label{sect:spectral}
Spectrally resolved indirect DM searches have the advantage of probing
different DM models through their characteristic spectral
distributions. To make current and future DM searches more effective it
is important to know in which energy band to focus the efforts in
order to maximize the potential DM signal over the expected
background. We dedicate this section to the spectral distribution of
gamma-ray flux from clusters. The flux emerge from a DM LP model that
emit final state radiation and give rise to substantial amounts of
electrons and positrons that IC upscatter background radiation fields
to high energies. We also consider a few supersymmetric DM models with
a high gamma-ray yield in the form of continuum emission and IC
induced emission. In addition to the annihilating DM, we estimate the
gamma-ray flux induced by shock accelerated CRps.

In Fig.~\ref{fig:diff_BM} we show the differential flux from the
Fornax cluster which is one of the best clusters for indirect DM
searches due to its relative high DM induced fluxes and low CR induced
fluxes. We derive the emission from four different supersymmetric BM
models and compare the flux to both the emission induced by CRps as
well as to the differential gamma-ray upper limits set by Fermi-LAT
after 18 months of observations \cite{2010ApJ...717L..71A}. We find
that the upper limits are not violated, although the predicted DM flux
that is dominated by the continuum emission from $\Kp$ and $\Ip$ models
(shown in the left panels) are close to the upper limits and can
potentially be probed by the Fermi 3 year data. Furthermore, the
expected masking gamma-ray signal induced by CRps is about a factor 10
below the DM continuum flux from the $\Kp$ and $\Ip$ BM models around
10~GeV. It is also important to note that IC emission from upscattered
CMB and SD photons is at least a factor 10 lower than the expected
flux from CRps above 100~MeV, making it very hard to distinguish such
a signal from the foreground due to the similar spectral
index. Furthermore, below the pion bump we expect the IC emission
coming from shock accelerated electrons to be dominating
\cite{2010MNRAS.409..449P}. Hence the IC emission from SUSY DM in
clusters can for most applications safely be neglected compared to the
CRps and DM continuum emission.

@@

CONSIDER REWRITING, START WITH PHYSICS, THEN DISCUSS DOMINATING
CONTRIBUTIONS, PEAKS, AND SPECTRAL INDICES

In Fig.~\ref{fig:IR_comp} we show the total IC emission as well as its
individual contributions from different IC upscattered radiation
fields. The left panel show the gamma-ray emission from the LP model
and the right panel from the $\Kp$ BM model, and for comparison we
overplot the emission expected from the CRps. Due to the flat electron
and positron spectrum resulting from the DM LP model and the relative
smaller energy density of the CMB compared to the SD, we find that the
upscattered CMB photons dominate the total DM IC emission in the
energy regime below 100~GeV, while the SD dominate above this
energy. For the BM models this transition is shifted towards a smaller
energy since the electron and positron spectrum has a steep spectral
distribution that suppresses low energy photon fields. This
suppression is also seen in the relative normalization of the flux
where the IC from both CMB and dust photons are lower compared to the
starlight simply because the starlight has a higher energy, hence it
is upscattered by more abundant low energy electrons and positrons
(c.f. Fig.~\ref{fig:q_DM}). At the highest gamma-ray energies of about
100~GeV and above, the IC from starlight steepens because it both
probes the high energy tail of the electrons as well as suffer from
the KN suppression. Even though the SD component seems to be
suppressed compared to the CMB at energies around 100~MeV, it has a
relative large impact on the steady state electron spectrum. If we
remove the SD cooling, we find a factor two larger IC flux at low
energies for a cluster of the size of Fornax. For more massive
clusters we expect a larger IC flux from the SD component, but a
smaller contribution from the CMB because of the more effective
cooling.

@@

It is also interesting to compare the total contribution from the LP
model, the brightest BM model, and the CRp induced emission. In
Fig.~\ref{fig:flux_int} we show the integrated flux from Fornax for
our different gamma-ray models and compare it to the unresolved
integrated flux upper limit on Fornax set by Fermi-LAT where they
averaged the flux over the energy range $0.2-100\,\gev$ assuming a
spectral index of 2. Again we are overproducing the upper limits,
although only with a factor 10 which is about a order magnitude less
constraining than the differential flux in the energy range
$1-10$~GeV. From the figure one clearly see that the LP model is
dominating the entire gamma-ray energy range up to the DM mass of this
model of about 1~TeV. In addition it is interesting to note that the
flux the DM BM $\Kp$ model is above the predicted emission from the
CRps in the $1-100$~GeV energy regime accessible both to Fermi and
Cherenkov telescopes.

@@

We continue by comparing how the estimated differential flux from
Fornax to three other clusters in Fig.~\ref{fig:clu_comp}; the close
by well studied Virgo cluster, the X-ray bright Perseus clusters, and
the massive merging Coma cluster. We find high DM induced gamma-ray
fluxes from both the Fornax and Virgo clusters, which highlight them
as promising targets for indirect DM searches. Note, however, that we
have used spatially unresolved upper limits, which is quite optimistic
for the angular extended Virgo cluster. It is quite striking how
constraining the upper limits for the Fornax cluster is compared to
the other clusters, especially in the $1-10$~GeV energy regime where
Fermi-LAT is most sensitive due to the maximized effective area. The
upper limits for Virgo and Perseus are background contaminated by AGN
activity from M87 and NGC 1267, respectively, and do not gain much
from the increased sensitivity. One should also note that the high CRp
induced flux from both Coma and Perseus is of the same magnitude as
the DM flux from LP models, making it difficult for indirect DM
searches in those clusters. Conversely is the Fornax cluster a great
target for indirect DM studies because the relative low gamma-ray flux
from CRps, absents of an active AGN, and high DM gamma-ray
flux. Actually, the Fermi-LAT 3 year data will be able to constrain
constrain the smallest mass of substructures using the DM BM
models. We have also learned that close by clusters are superior to
far away clusters, even though close by clusters have a mass that is
much smaller. In fact, we can understand this by looking at the
luminosity-to-mass scaling relations. The gamma-ray luminosity from
the smooth density profile is given by
\begin{equation}
L_\rmn{sm} \propto \int \dd V \rho(r)^2 \propto \frac{c^3}
{\left[\log\left(1+c\right)-c/(1+c)\right]} \propto \mvir^{0.83}\,,
\end{equation}
and the total luminosity including boostfactors for the DM LP and BM
models is given by
\begin{eqnarray}
L_\rmn{tot,LP} &=& L_\rmn{sm,LP} \B_\rmn{LP} \propto \mvir^{0.73}\qquad
\,\,\,\,\,\,\rmn{and}\\
L_\rmn{tot,BM} &=& L_\rmn{sm,BM} \B_\rmn{BM} \propto  \mvir^{1.06}\,,\qquad
\rmn{where} \\
B_\rmn{LP}&=&B_\sfe B_\sub \propto \mvir^{-1/3}\mvir^{0.23} \propto
 \mvir^{-0.1}\,,\\
B_\rmn{BM}&=&B_\sub \propto \mvir^{0.23}\,.
\end{eqnarray}
Hence increasing the mass with a factor two doubles the luminosity,
but increasing the distance by a factor two decrease the flux with a
factor four.
\del{
\begin{eqnarray}
L_{\gamma,\rmn{sub tot}} \propto N\,L_{\gamma,\sub} \propto N\,M_\sub 
 \propto M_\sub^{\left(A+1\right)}\,\frac{\dd N_\sub}{\dd M} \propto  M^{A-0.9}
\end{eqnarray}
Assuming that the concentration $c$ decrease for increasing mass as it
does for the mass range $10^{12}-10^{14}\,\msun$ where
$L_{\gamma,\sub} \propto N\,M_\sub^A$, where $A<1$.
}

>>>>

{\it IC -- dust and starlight}\\ Galaxy clusters are typically
characterized by the hot gas in the ICM and the collection of
gravitationally bound galaxies. The emission in the IR and UV
wavelengths emerge from dust and starlight in both the galaxies and
the ICM (e.g. \cite{2006ApJ...648L..29P} and
\cite{2009MNRAS.399.1694G}). Much of the radiation leaks out from the
galaxy into the ICM, which give rise to similar spectral distributions
for the galaxies and the ICM in these wavelengths. Hence, we use the
accurately measured light from galactic dust and starlight for the ISM
to model the ICM. We characterize the spectra through a fit to the
galactic spectra presented in \cite{2006ApJ...648L..29P}. The spatial
distribution of the SD light in clusters is extracted from a stacked
emission analysis of Sloan Digital Sky Survey (SDSS) data at the
redshift $\sim 0.25$ \cite{2005MNRAS.358..949Z}. Finally, we fix the
normalization for the light from stars and dust individually in a
galaxy cluster using using the IR to X-ray luminosity scaling relation
found in \cite{2008A&A...490..547G}. The X-ray luminosities for the
clusters used in this work are listed in the extended HIghest X-ray
FLUx Galaxy Cluster Sample (HIFLUGCS) \cite{2002ApJ...567..716R} from
the Rosat all-sky survey. Note that we have added the Virgo cluster to
the sample that we nevertheless refer to as extended HIFLUGCS
catalogue in the following.

{\it Spatial distribution -- stars and dust}\\ Our goal in this
section is to derive a simple spatial model for the distribution of SD
in galaxy clusters. For this reason we use stacked cluster
observations by Sloan in the $r$-band that trace the stars. We assume
that the dust trace the stars in the clusters, which should be
accurate to the order of magnitude estimate that we are interested in
here. The stacked surface brightness profiles in
\cite{2005MNRAS.358..949Z} are measured at redshift $z \sim 0.25$ with
an average mass of the clusters of $4.0\times10^{13}\,\msun$. The
profiles are build up out of three components; diffuse intracluster
light (ICL), galaxies, and the brightest cluster galaxy (BCG) in the
center of clusters. However, the total volume overlap of the ISM to
the cluster is very small compared to the ICL- and BCG overlap with
the cluster, thus the relative contribution of the galaxies to the IC
emission is suppressed. To remove this bias we cut out the galaxies
that have been smoothed over the cluster in the stacked analysis. In
Fig.~\ref{fig:SD_spatial} we show the SDSS stacked brightness profiles
as well as the fitted profiles\footnote{The measured brightness is
  converted into units of $\rmn{erg}\,\rmn{s}^{-1}$ using
  \cite{2010...book} $S(\rmn{mag}\,''^{-2}) =
  \mathcal{M}_\odot+21.572-2.5\rmn{log}_{10}S(L_\odot\,\rmn{pc}^{-2})$,
  where the sun's absolute magnitude in the $r$-band is given by
  $\mathcal{M}_\odot=27.1$ \cite{2000asqu.book..339L} and the
  luminosity of the sun by $L_\odot=3.85\,10^{33}
  \rmn{erg}\,\rmn{s}^{-1}$.}, both scaled with the unitless $\rvir$
normalized area to resembele the expected luminosity at radius
$r$. Our benchmark spatial model is shown with the solid black line,
and include the contribution from the ICL and BCG.

@@

In Fig.~\ref{fig:SD_spectra} we show the spectral distribution of the
light from SD together with the CMB black-body distribution in a
galaxy cluster, both at the radius $r=0.071\rvir$ where the energy
density of the smooth light from SD (see Fig.~\ref{fig:SD_Edens})
equals the energy density of the CMB. We have renormalized both the
spectral data and the fitted spectra using the luminosity from stars
and dust individually (see the normalization section below for further
details).
\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/fit.porter.v2.eps}
\caption{\it Spectral dependence of radiation fields in a cluster of
  galaxies. The black dotted line in the left peak show the spectrum
  of CMB photons using a black body with a temperature of
  $2.73$~K. The crosses in the middle and right peaks represent the
  measured spectra from stars and dust (SD), respectively, and are
  derived in \cite{2006ApJ...648L..29P} for a galaxy. We normalize the
  individual SD spectrum separately using the observed luminosity from
  SD in clusters. The SD luminosity is related to the cluster mass
  through Eq, where we use have used the mass
  $\mvir=4.0\times10^{13}\msun$ in this figure. Finally we renormalize
  the SD spectra to the radius $r=0.071\rvir$, where the smooth energy
  density of the SD light (see Fig.~\ref{fig:SD_Edens}) equals the
  energy density of the CMB. The red dashed lines show the fitted SD
  spectral model.}
 \label{fig:SD_spectra}
\end{figure}
We choose to fit the spectrum of the dust that peaks at about
$10^{-2}\,\ev$ with a double power-law, and the slightly wider
spectral distribution of the stars that peaks at about $1\,\ev$ with a
tipple power-law. We find the best fit spectra for the SD in a
cluster through:
\begin{eqnarray}
  \eph^2\frac{\dd^2 N_\ph}{\dd \eph \dd V}
  &\equiv& u_\sd(\eph) =  \sum_i N_i(\lx) u_i^\gal(\eph)\,,\nonumber \\ 
\rmn{where}\quad i&=&\{\rmn{stars,dust}\} \qquad \rmn{and}\\ 
% &=&  N_\dust(\lx) u_\dust^\gal(\eph) + N_\stars(\lx) u_\stars^\gal(\eph)  \nonumber \\ 
%  \\
  u_\stars^\gal(\eph) &=& \frac{23\,\rmn{eV}}{\rmn{cm}^3} 
  \left(\frac{1.23\,\rmn{eV}}{\eph}\right)^{1.9} \nonumber \\
  &\times&\left[1+\left(\frac{2.04\,\rmn{eV}}{\eph}\right)^{20}\right]
  ^{-\frac{1.9}{20}}\nonumber \\
  &\times& \left[1+\left(\frac{0.78\,\rmn{eV}}{\eph}\right)^{20}\right]^{-\frac{2.6}{20}}\,, \\
  u_\dust^\gal(\eph) &=& 
  \frac{40\,\rmn{eV}}{\rmn{cm}^3} 
  \left(\frac{0.0144\,\rmn{eV}}{\eph}\right)^{4.9}\nonumber \\
  &\times& \left[1+\left(\frac{0.0144\,\rmn{eV}}{\eph}\right)^{1.9}\right]^{-4.9}\,.
\end{eqnarray}

{\it Normalization}\\ The spatial and spectral distribution functions
of the light from SD in a cluster are both normalized with a constant
determined from the luminosity in each respective wavelength. The
specific energy density is given by
\begin{eqnarray} 
u_\sd(\eph, r) &=& \frac{j(r)}{j(0)}\,u_\sd(\eph) \nonumber\\
&=& \frac{j(r)}{j(0)}\,\sum_i N_i(\lx) u_i^\gal(\eph)\,,
\end{eqnarray}
where we fix the unitless normalization $N_i(\lx)$ using the total
energy of the light in SD  within $\rvir$:
\begin{eqnarray} 
  E_{i,\vir} &=& L_i \frac{\rvir}{c} \nonumber \\
  &=&N_i(\lx)\int_{\rvir} \int_i \,\frac{j(r)}{j(0)} 
  \frac{u_i^\gal(\eph)}{\eph}\,\dd V\dd \eph\,,\nonumber \\
 \rmn{for}\quad i&=&\{\rmn{stars,dust}\}\,.
\label{eq:E_SD}
\end{eqnarray}
Here we approximate the total energy of the photons within $\rvir$
($E_{i,\vir}$) with the luminosity $L_i$ \cite{2008A&A...490..547G}
times the typical timescale it takes for a photon to propagate through
the cluster (we assume that the cluster is optically thin). Once we
derived the SD luminosities for a well representative cluster, we use
a IR to X-ray scaling relation to estimate the luminosity from an
arbitrary cluster:
\begin{equation}
L_\ir=1.05\times10^{45}\,\rmn{erg\,s}^{-1}\,
\left(\frac{\lx}{10^{44}\rmn{erg\,s}^{-1}}\right)^{0.53}\,,
\label{eq:SD-X}
\end{equation}
where $\lx$ is the luminosity of the X-ray emitting gas within
$\rvir$.

@@

{\it Energy density}\\ The spatial distribution of the SD energy
density is an important quantity since it impacts the steady state
electron and positron spectra through cooling. A higher density of SD
in a cluster suppresses the spectra of the upscattering particles. With
the formalism presented above, we can now derive the energy density
from starlight and dust in a galaxy cluster through
\begin{eqnarray}
\label{eq:U_SD}
u_\sd(r) &=& \int \dd \eph \frac{\dd^2 N_\ph}{\dd \eph \dd V}\,\eph
=\int \dd \eph \frac{u_\sd(\eph, r)}{\eph}
\nonumber \\
&=&  \frac{j(r)}{j(0)}  \int \dd \eph \sum_i 
N_i(\lx) \frac{u_i^\gal(\eph)}{\eph}\,. \nonumber \\
\end{eqnarray}
In Fig.~\ref{fig:SD_Edens} we compare the energy densities from
different radiation fields in a galaxy cluster with the mass
$\mvir=4.0\times10^{13}\msun$. We use two different profiles for the
SD energy density, where the total profile includes the contribution
from the ICL, the BCG and all the galaxies, while the galaxies are
excluded in the smooth profile. We find that even for a cluster with a
relative small mass the energy density of the SD components dominates
over the CMB and the magnetic fields (with a central B field of 3~$\mu
G$) within about 10\% of $\rvir$. Outside this radius the CMB is
dominating the energy density of the cluster, hence the magnetic
fields are always subdominant in the cluster and can safely be removed
from cluster IC calculations for a reasonable central B-field and
spectral index. Note that in this work we keep the contribution
from B-fields to the total energy density for consistency.

@@

\section{Spectral gamma-ray distribution}
\label{sect:spectral}
Spectrally resolved indirect DM searches have the advantage of probing
different DM models through their characteristic spectral
distributions. To make current and future DM searches more effective
it is important to know in which energy band to focus the efforts in
order to maximize potential DM signal over expected background.

We dedicate this section to the spectral distribution of gamma-ray
flux from clusters. The flux emerge from a leptophilic DM model that
emit final state radiation and give rise to substantial amounts of
electrons and positrons that IC upscatter background radiation fields
to high energies. We also consider a few supersymmetric DM models with
a high gamma-ray yield in the form of continuum emission and IC
induced emission. In addition to the annihilating DM, we estimate the
gamma-ray flux induced by shock accelerated CRps.

In Fig.~\ref{fig:diff_BM} we show the differential flux from the
Fornax cluster which is one of the best clusters for indirect DM
searches due to its relative high DM induced fluxes and low CR induced
fluxes. We derive the emission from four different supersymmetric BM
models and compare the flux to both the emission induced by CRps as
well as to the differential gamma-ray upper limits set by Fermi-LAT
after 18 months of observations \cite{2010ApJ...717L..71A}. We find
that the upper limits are not violated, although the predicted DM flux
that is dominated by the continuum emission from $\Kp$ and $\Ip$
models (shown in the left panels) are close to the upper limits and
can potentially be probed by the Fermi 3 year data. Furthermore, the
expected masking gamma-ray signal induced by CRps is about a factor 10
below the DM continuum flux from the $\Kp$ and $\Ip$ BM models around
10~GeV. However, the IC emission from upscattered CMB and SD photons
is at least a factor 10 lower than the expected flux from CRps above
100~MeV, making it very hard to distinguish such a signal from the
foreground due to the similar spectral index. For energies below
100~MeV we expect shock accelerated electrons and positrons to be
dominating \cite{2010MNRAS.409..449P} over the supersymmetric DM
induced leptons. Hence the IC emission from supersymmetric DM in
clusters can for most applications safely be neglected compared to the
CRps and DM continuum emission.

@@

\del{CONSIDER REWRITING, START WITH PHYSICS, THEN DISCUSS DOMINATING
CONTRIBUTIONS, PEAKS, AND SPECTRAL INDICES}

In Fig.~\ref{fig:IR_comp} we show the total IC emission as well as its
individual contributions from different IC upscattered radiation
fields. The left panel show the gamma-ray emission from the LP model
and the right panel from the $\Kp$ BM model, and for comparison we
overplot the emission expected from the CRps. Due to the flat electron
and positron spectrum resulting from the LP DM model and the smaller
mean energy of the CMB compared to the SD, we find that the
upscattered CMB photons dominate the total DM IC emission in the
energy regime below 100~GeV, while the SD dominate above this
energy. For the BM models this transition is shifted towards a smaller
energy since the electron and positron spectrum has a steep spectral
distribution that suppresses low energy photon fields. This
suppression is also seen in the relative normalization of the flux
where the IC from both CMB and dust photons are lower compared to the
starlight simply because the starlight has a higher energy, hence it
is upscattered by more abundant low energy electrons and positrons
(c.f. Fig.~\ref{fig:q_DM}). At the highest gamma-ray energies of about
100~GeV and above, the IC from starlight steepens because it both
probes the high energy tail of the electrons as well as suffer from
the KN suppression. Even though the SD component seems to be
suppressed compared to the CMB at energies around 100~MeV, it has a
relative large impact on the steady state electron spectrum. If we
remove the SD cooling, we find a factor two larger IC flux at low
energies for a cluster of the size of Fornax. For more massive
clusters we expect a larger IC flux from the SD component, but a
smaller contribution from the CMB because of the more effective
cooling.

@@

It is also interesting to compare the total contribution from the LP
model, the brightest BM model ($\Kp$), and the CRp induced emission. In
Fig.~\ref{fig:flux_int} we show the integrated flux from Fornax for
our different gamma-ray models and compare it to the unresolved
integrated flux upper limit on Fornax set by Fermi-LAT where they
averaged the flux over the energy range $0.2-100\,\gev$ assuming a
spectral index of 2. Again we are overproducing the upper limits,
although only with a factor 10 which is about a order magnitude less
constraining than the differential flux in the energy range
$1-10$~GeV. From the figure one clearly see that the LP model is
dominating the entire gamma-ray energy range up to the DM mass of this
model of about 1~TeV. In addition it is interesting to note that the
flux the DM BM $\Kp$ model is above the predicted emission from the
CRps in the $1-100$~GeV energy regime accessible both to Fermi and
Cherenkov telescopes.

@@

We continue by comparing the estimated differential flux from Fornax
to three other clusters in Fig.~\ref{fig:clu_comp}; the close by well
studied Virgo cluster, the X-ray bright Perseus clusters, and the
massive merging Coma cluster. We find high DM induced gamma-ray fluxes
from both the Fornax and Virgo clusters, which highlight them as
promising targets for indirect DM searches. Note, however, that we
have used spatially unresolved upper limits, which is quite optimistic
for the angular extended Virgo cluster. It is quite striking how
constraining the upper limits for the Fornax cluster is compared to
the other clusters, especially in the $1-10$~GeV energy regime where
Fermi-LAT is most sensitive due to the maximized effective area. The
upper limits for Virgo and Perseus are background contaminated by AGN
activity from M87 and NGC 1267, respectively, and do not gain much
from the increased sensitivity. One should also note that the high CRp
induced flux from both Coma and Perseus is of the same magnitude as
the DM flux from LP models, making it difficult for indirect DM
searches in those clusters. Conversely is the Fornax cluster a great
target for indirect DM studies because the relative low gamma-ray flux
from CRps, absents of an active AGN, and high DM gamma-ray
flux. Actually, the Fermi-LAT 3 year data will be able to constrain
the smallest mass of substructures using the DM BM models. We have
also learned that the closest clusters are superior for DM searches
compared to further away clusters, even though the close by clusters
have a mass that is smaller. In fact, we can understand this by
looking at the luminosity-to-mass scaling relations. The gamma-ray
luminosity from the smooth density profile is given by
\begin{equation}
L_{\gamma,\rmn{sm}} \propto \int \dd V \rho(r)^2 \propto \frac{c^3}
{\left[\log\left(1+c\right)-c/(1+c)\right]} \propto \mvir^{0.83}\,,
\end{equation}
and the total DM luminosity including boost factors for the LP and
the BM models are given by
\begin{eqnarray}
L_{\gamma,\rmn{LP}} &=& L_{\gamma,\rmn{smLP}} \B_\rmn{LP} \propto \mvir^{0.73}\qquad
\,\,\,\,\,\,\rmn{and}\\
L_{\gamma,\rmn{BM}} &=& L_{\gamma,\rmn{smBM}} \B_\rmn{BM} \propto  \mvir^{1.06}\,,\qquad
\rmn{where} \\
B_\rmn{LP}&=&B_\sfe B_\sub \propto \mvir^{-1/3}\mvir^{0.23} \propto
 \mvir^{-0.1}\,,\\
B_\rmn{BM}&=&B_\sub \propto \mvir^{0.23}\,.
\end{eqnarray}
Hence increasing the mass with a factor two doubles the luminosity,
but increasing the distance by a factor two decrease the flux with a
factor four.
\del{
\begin{eqnarray}
L_{\gamma,\rmn{sub tot}} \propto N\,L_{\gamma,\sub} \propto N\,M_\sub 
 \propto M_\sub^{\left(A+1\right)}\,\frac{\dd N_\sub}{\dd M} \propto  M^{A-0.9}
\end{eqnarray}
Assuming that the concentration $c$ decrease for increasing mass as it
does for the mass range $10^{12}-10^{14}\,\msun$ where
$L_{\gamma,\sub} \propto N\,M_\sub^A$, where $A<1$.
}
